---
title: "Space Compost - Ammending Regolith"
authors: "Sedona Spann"
date: "8 April, 2026"
format: html
editor: visual
---

# Load Libraries

```{r}
#| message: false
#| warning: false
library(readxl)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(emmeans)
library(viridis)   # For colorblind-safe palette
library(ggbeeswarm) # Optional: for nice jittered points
library(sjPlot)
library(ggeffects)
library(ggsignif)
library(performance)
library(lme4)
library(nlme)
library(ggpubr)
library(glmmTMB)
library(multcomp)
library(multcompView)
library(rstatix)
library(broom)
library(ggh4x)
library(MASS)
library(tidyverse)
library(betareg)
library(gamlss)
library(caret)
library(DHARMa)
```

# Load Data Clean and Create palettes

```{r}
#| message: false
#| warning: false
ds_nu <- read_excel("C:/Users/sms689/Downloads/space-compost-metadata-2.2026.xlsx",
                 sheet = "magentabox_metadata") %>%
                 clean_names() %>%
                 mutate(total_biomass_mg = overall_biomass_g * 1000) %>%
                 mutate(shoot_biomass_mg = shoot_biomass_dry_g * 1000) %>%
                 mutate(hec = factor(hec, levels = c("sterile","live","none")),
                        amf = factor(amf, levels = c("live", "sterile")))%>%
                 mutate(root_biomass_mg = root_biomass_dry_g * 1000)%>%
                 mutate(root_shoot_ratio_mg = root_biomass_mg / shoot_biomass_mg)%>%
                 mutate(total_chl = spad_avg * shoot_biomass_dry_g)%>%
                 filter(!is.na(total_biomass_mg) ) %>%
                 filter(urea == "none")
#View(ds_nu)

# Reference mean: mean of sterile-amf within each hec group
# (used for both hec_gr and amf_gr)
amf_sterile_means <- ds_nu %>%
  filter(amf == "sterile") %>%
  group_by(hec) %>%
  summarise(amf_sterile_mean = mean(total_biomass_mg, na.rm = TRUE), .groups = "drop") %>%
  pull(amf_sterile_mean)

hec_none_means <- ds_nu %>%
  filter(hec == "none") %>%
  group_by(amf) %>%
  summarise(hec_none_mean = mean(total_biomass_mg, na.rm = TRUE), .groups = "drop") %>%
  pull(hec_none_mean)

ref_means <- ds_nu %>%
  filter(hec == "none", amf == "sterile") %>%
  summarise(ref_mean = mean(total_biomass_mg, na.rm = TRUE)) %>%
  pull(ref_mean)

#View(ref_means) 14.1266666666667

ds_nu <- ds_nu %>%
  mutate(
    hec_gr = if_else(
      hec %in% c("live", "sterile"),
      log(total_biomass_mg / hec_none_means),
      NA_real_
    ),
    amf_gr = if_else(
      amf == "live",
      log(total_biomass_mg / amf_sterile_means),
      NA_real_
    ),
    gr =log(total_biomass_mg / ref_means)
    )

write.csv(ds_nu, "C:/Users/sms689/Downloads/ds_nu.csv", row.names = FALSE)
#ds_nu <- read.csv("C:/Users/sms689/Downloads/ds_nu.csv")

pal_hec <- c("none" = "#818186", "live" = "#73b343", "sterile" = "#7f4753")

pal_amf <- c("live" = "#5e8dbb", "sterile" = "#b56060")
```

# Checking the structure of data and normality of response variable
```{r}
num_vars <- ds_nu %>%
  dplyr::select(where(is.numeric)) %>%
  names()

ds_nu %>%
  dplyr::select(all_of(num_vars)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  ) %>%
  ggplot(aes(x = value)) +
  geom_histogram(
    bins = 30,
    fill = "gray70",
    color = "black"
  ) +
  facet_wrap(~ variable, scales = "free") +
  theme_bw(base_size = 11) +
  theme(
    strip.text = element_text(face = "bold")
  ) +
  labs(
    x = NULL,
    y = "Count",
    title = "Histograms of Continuous Variables (Normality Check)"
  )
## The histogram of overall biomass looks right skewed  
```

## Boxplot of Total Biomass by Replicate

```{r}
mutate(ds_nu, replicate = as.factor(replicate)) %>%
ggplot(aes(x = replicate, y = overall_biomass_g, fill = replicate)) +
  geom_boxplot()
```

## Model Selection

```{r}
random_m <- lmer(overall_biomass_g ~ amf * hec + (1|replicate), ds_nu)
interactions_m <- lm(overall_biomass_g ~  amf * hec, ds_nu)
additive_m <- lm(overall_biomass_g ~  amf + hec, ds_nu)
AIC(random_m, interactions_m, additive_m)
DHARMa::testResiduals(additive_m)
DHARMa::testResiduals(interactions_m)
DHARMa::testResiduals(random_m)
anova(interactions_m)
anova(additive_m)

random_int_m <- lmer(spad_avg ~ amf * hec + (1|replicate), ds_nu)
random_add_m <- lmer(spad_avg ~ amf + hec + (1|replicate), ds_nu)
interactions_m <- lm(spad_avg ~  amf * hec, ds_nu)
additive_m <- lm(spad_avg ~  amf + hec, ds_nu)
AIC(random_int_m, random_add_m, interactions_m, additive_m)

ggplot(ds_nu, aes(x = amf, y = spad_avg, fill = hec)) +
  geom_boxplot() +
  theme_bw()

colnames(ds_nu)
```

# Plots

## Reusable boxplot function

```{r}
# ---- Reusable boxplot function ----
plot_boxplot_cld <- function(data, response_var, x_var, fill_var,
                              palette,
                              x_label, fill_label, y_label, title, 
                              hline = NULL,
                              caption_extra = NULL,
                              ymax_offset = 0.5,
                              facet_var = NULL,
                              facet_labeller = NULL) {
  
  # 1. Fit separate models per facet if facet_var is specified, otherwise fit one model
if (!is.null(facet_var)) {
  
  split_data <- data %>%
    group_by(.data[[facet_var]]) %>%
    group_split()
  
  facet_levels <- data %>%
    distinct(.data[[facet_var]]) %>%
    pull()

  cld_list <- lapply(seq_along(split_data), function(i) {
    df <- split_data[[i]]
    facet_value <- facet_levels[i]
    
    formula_mod <- as.formula(paste(response_var, "~", fill_var, "*", x_var))
    formula_emm <- as.formula(paste("~", fill_var, "|", x_var))
    
    mod <- lm(formula_mod, data = df)
    emm <- emmeans(mod, formula_emm)
    
    multcomp::cld(emm, Letters = letters) %>%
      as.data.frame() %>%
      mutate(
        .group = trimws(.group),
        !!facet_var := facet_value
      )
  })
  
  cld_df <- bind_rows(cld_list)

} else {

  formula_mod <- as.formula(paste(response_var, "~", fill_var, "*", x_var))
  formula_emm <- as.formula(paste("~", fill_var, "|", x_var))
  
  mod <- lm(formula_mod, data = data)
  emm <- emmeans(mod, formula_emm)
  # 2. CLD
  cld_df <- multcomp::cld(emm, Letters = letters) %>%
    as.data.frame() %>%
    mutate(.group = trimws(.group))
}
  # 3. Whisker tops
  group_vars <- c(x_var, fill_var, if (!is.null(facet_var)) facet_var)
  
  whisker_df <- data %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(ymax = boxplot.stats(.data[[response_var]])$stats[5],
              .groups = "drop")
  
  cld_df <- cld_df %>%
    left_join(whisker_df, by = group_vars)
  
  # 4. Plot
  p <- ggplot(data, aes(x = .data[[x_var]], y = .data[[response_var]],
                        fill = .data[[fill_var]])) +
    geom_boxplot(outlier.shape = NA, alpha = 0.85, color = "gray30",
               stat = "summary",
               fun.data = function(x) {
                 m   <- mean(x, na.rm = TRUE)
                 sem <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
                 data.frame(
                   ymin   = min(x, na.rm = TRUE),  # whisker bottom
                   lower  = m - sem,
                   middle = m,
                   upper  = m + sem,
                   ymax   = max(x, na.rm = TRUE)   # whisker top
                 )
               }
  ) +
    geom_point(
      aes(color = .data[[fill_var]]),
      position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
      alpha = 0.5, size = 2
    ) +
    { if (!is.null(hline)) geom_hline(yintercept = hline, linetype = "dashed", color = "gray40") } +
    scale_fill_manual(name = fill_label, values = palette) +
    scale_color_manual(name = fill_label, values = palette) +
    geom_text(
      data = cld_df,
      aes(x = .data[[x_var]], y = ymax + ymax_offset,
          label = .group, group = .data[[fill_var]]),
      position = position_dodge(width = 0.75),
      size = 4.5, inherit.aes = FALSE
    ) +
    theme_bw(base_size = 12) +
    theme(
      plot.title       = element_text(face = "bold", hjust = 0.5, size = 18),
      axis.title       = element_text(face = "bold", size = 15),
      axis.text.x      = element_text(size = 15),
      axis.text.y      = element_text(size = 15),
      panel.grid.minor = element_blank(),
      panel.border     = element_rect(color = "black", fill = NA),
      strip.text       = element_text(face = "bold")
    ) +
    labs(
      y = y_label, x = x_label, title = title,
      subtitle = "Mean ± SEM, n = 3 per condition",
      caption = paste("Letters indicate significant differences within HEC treatments. Linear Model:",
                      response_var, "~", x_var, "*", fill_var,
                      if (!is.null(caption_extra)) paste("\n", caption_extra) else "")
    ) 

  # 5. Add facet if specified
  if (!is.null(facet_var)) {
    p <- p + facet_wrap(
      as.formula(paste("~", facet_var)),
      labeller = if (!is.null(facet_labeller)) facet_labeller else label_value
    )
  }
  
  p
}
```

## total biomass boxplot hec on x-axis, amf as fill

```{r}
#### Total Biomass by HEC and AMF Treatment
#| fig-width: 12
#| fig-height: 8
plot_boxplot_cld(
  data         =  ds_nu,
  response_var = "total_biomass_mg",      
  x_var        = "hec",
  fill_var     = "amf",
  palette      =  pal_amf,
  x_label      = "HEC Treatment",
  fill_label   = "AMF Treatment",
  y_label      = "Total Biomass (mg)",
  title        = "Total Biomass by HEC and AMF Treatment",
  ymax_offset  = 5                     # adjust letter height per variable
)
ggsave("C:\\Users\\sms689\\Downloads\\total_biomass_hec_fill_with_amf_boxplot.bmp", width = 8, height = 6, dpi = 300)
```

## Total Biomass Boxplot with AMF on x-axis and HEC as fill, with CLD letters for significant differences

```{r}
#| fig-width: 12
#| fig-height: 8

# 1. Model and emmeans (no urea since this is ds_nu)
mod <- lm((overall_biomass_g*1000) ~ amf * hec, data = ds_nu)
emm <- emmeans(mod, ~ amf | hec)

# 2. CLD — cld_df already has amf and hec from emmeans, no join needed
cld_df <- multcomp::cld(emm, Letters = letters) %>%
  as.data.frame() %>%
  mutate(.group = trimws(.group))

# 3. Get whisker tops to position CLD letters
whisker_df <- ds_nu %>%
  group_by(amf, hec) %>%
  summarise(ymax = boxplot.stats(overall_biomass_g*1000)$stats[5], .groups = "drop")

cld_df <- cld_df %>%
  left_join(whisker_df, by = c("amf", "hec"))

# 4. Plot
ds_nu %>%
  ggplot(aes(x = amf, y = overall_biomass_g*1000, fill = hec)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.85, color = "gray30",
             stat = "summary",
             fun.data = function(x) {
               data.frame(
                 ymin   = min(x),
                 lower  = quantile(x, 0.25),
                 middle = mean(x),          # mean instead of median
                 upper  = quantile(x, 0.75),
                 ymax   = max(x)
               )
             }) +
  geom_point(
    aes(color = hec),
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
    alpha = 0.5,
    size = 2
  ) +
  scale_fill_manual(name = "HEC Treatment", values = pal_hec) +
  scale_color_manual(name = "HEC Treatment", values = pal_hec) +
  geom_text(
    data = cld_df,
    aes(x = amf, y = ymax + 5, label = .group, group = hec),
    position = position_dodge(width = 0.75),
    size = 4.5,
    inherit.aes = FALSE
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", hjust = 0.5, size = 18),
    axis.title       = element_text(face = "bold", size = 15),
    axis.text.x      = element_text(size = 15),
    axis.text.y      = element_text(size = 15),
    panel.grid.minor = element_blank(),
    panel.border     = element_rect(color = "black", fill = NA),
    strip.text       = element_text(face = "bold")
  ) +
  labs(
    y = "Total Biomass (mg)",
    x = "AMF Treatment",
    title = "Total Biomass by AMF and HEC Treatment",
    caption = "Letters indicate groups that are significantly different from each other. Results based on Total Biomass ~ AMF × HEC"
  )
ggsave("C:\\Users\\sms689\\Downloads\\total_biomass_amf_fill_with_hec_boxplot.bmp", width = 8, height = 6, dpi = 300)
```

## root biomass boxplot hec on x-axis, amf as fill
```{r}
plot_boxplot_cld(
  data         =  ds_nu,
  response_var = "root_biomass_mg",      
  x_var        = "hec",
  fill_var     = "amf",
  palette      =  pal_amf,
  x_label      = "HEC Treatment",
  fill_label   = "AMF Treatment",
  y_label      = "Root Biomass (mg)",
  title        = "Root Biomass by HEC and AMF Treatment",
  ymax_offset  = 1                     # adjust letter height per variable
)
```

## shoot biomass boxplot hec on x-axis, amf as fill
```{r}
plot_boxplot_cld(
  data         =  ds_nu,
  response_var = "shoot_biomass_mg",      
  x_var        = "hec",
  fill_var     = "amf",
  palette      =  pal_amf,
  x_label      = "HEC Treatment",
  fill_label   = "AMF Treatment",
  y_label      = "Shoot Biomass (mg)",
  title        = "Shoot Biomass by HEC and AMF Treatment",
  ymax_offset  = 5                     # adjust letter height per variable
)
```

## spad boxplot hec on x-axis, amf as fill
```{r}
## Spad by HEC and AMF Treatment
#| fig-width: 12
#| fig-height: 8
plot_boxplot_cld(
  data         =  ds_nu,
  response_var = "spad_avg",
  x_var        = "hec",
  fill_var     = "amf",
  palette      =  pal_amf,
  x_label      = "HEC Treatment",
  fill_label   = "AMF Treatment",
  y_label      = "Relative Leaf Chlorophyll Content",
  title        = "Relative Leaf Chlorophyll Content (SPAD) \n by HEC and AMF Treatment",
  ymax_offset  = 0.7                     # adjust letter height per variable
)
ggsave("C:\\Users\\sms689\\Downloads\\Chlorophyll_boxplot.bmp", width = 9, height = 5, dpi = 300)

shell.exec("C:\\Users\\sms689\\Downloads\\Chlorophyll_boxplot.bmp")
```

## total chlorophyll boxplot hec on x-axis, amf as fill
```{r}
## total_chl by HEC and AMF Treatment
#| fig-width: 12
#| fig-height: 8
plot_boxplot_cld(
  data         =  ds_nu,
  response_var = "total_chl",
  x_var        = "hec",
  fill_var     = "amf",
  palette      =  pal_amf,
  x_label      = "HEC Treatment",
  fill_label   = "AMF Treatment",
  y_label      = "Total Relative Leaf Chlorophyll Content",
  title        = "Total Relative Leaf Chlorophyll Content (SPAD x Shoot Biomass grams) \n by HEC and AMF Treatment",
  ymax_offset  = 0.1                     # adjust letter height per variable
)

ggsave("C:\\Users\\sms689\\Downloads\\Total_Chlorophyll_boxplot.bmp", width = 10, height = 6, dpi = 300)

shell.exec("C:\\Users\\sms689\\Downloads\\Total_Chlorophyll_boxplot.bmp")
```

## root to shoot ratio boxplot hec on x-axis, amf as fill
```{r}

## root_shoot ratio by HEC and AMF Treatment
#| fig-width: 12
#| fig-height: 8
plot_boxplot_cld(
  data         =  ds_nu,
  response_var = "root_shoot_ratio_mg",
  x_var        = "hec",
  fill_var     = "amf",
  palette      =  pal_amf,
  x_label      = "HEC Treatment",
  fill_label   = "AMF Treatment",
  y_label      = "Root to Shoot Ratio (mg)",
  title        = "Root to Shoot Ratio by HEC and AMF Treatment",
  ymax_offset  = 0.1                       # adjust letter height per variable
)
ggsave("C:\\Users\\sms689\\Downloads\\Root_to_Shoot_Ratio_boxplot.bmp", width = 9, height = 8, dpi = 300)
```

## root alongside shoot biomass boxplot on one graph, hec on x-axis, amf as fill
```{r}
ds_long <- ds_nu %>%
  pivot_longer(
    cols = c(root_biomass_mg, shoot_biomass_mg),
    names_to = "tissue",
    values_to = "biomass_mg"
  ) %>%
  mutate(tissue = recode(tissue,
    "root_biomass_mg"  = "Root",
    "shoot_biomass_mg" = "Shoot"
  ))

plot_boxplot_cld(
  data         = ds_long,
  response_var = "biomass_mg",
  x_var        = "hec",
  fill_var     = "amf",
  facet_var    = "tissue",
  palette      = pal_amf,
  x_label      = "HEC Treatment",
  fill_label   = "AMF Treatment",
  y_label      = "Biomass (mg)",
  title        = "Root vs. Shoot Biomass by HEC and AMF Treatment",
  ymax_offset  = 5
)
ggsave("C:\\Users\\sms689\\Downloads\\Root_alongside_Shoot_Biomass_boxplot.bmp", width = 8, height = 6, dpi = 300)
```


## scatter plot of biomass vs spad with regression line and confidence interval, colored by hec
```{r}
cor.test(ds_nu$shoot_biomass_mg, ds_nu$spad_avg, method = "pearson")

ggplot(ds_nu, aes(x = shoot_biomass_mg, y = spad_avg)) +
  geom_point(aes(color = hec)) +
  facet_wrap(~ amf) +
  scale_color_manual(name = "HEC Treatment", values = pal_hec) +
  theme_bw() +
  geom_smooth(method = "lm") +
  labs(x = "Total Biomass (g)", y = "SPAD", 
       title = "Chlorophyll Dilution: SPAD vs Biomass")
```

## scatter plot of biomass vs spad with regression line and confidence interval, colored by amf
```{r}
cor.test(ds_nu$shoot_biomass_mg, ds_nu$spad_avg, method = "pearson")

ggplot(ds_nu, aes(x = shoot_biomass_mg, y = spad_avg)) +
  geom_point(aes(color = amf)) +
  scale_color_manual(name = "AMF Treatment", values = pal_amf) +
  facet_wrap(~ hec) +
  theme_bw() +
  geom_smooth(method = "lm") +
  labs(x = "Total Biomass (g)", y = "SPAD", 
       title = "Chlorophyll Dilution: SPAD vs Biomass")

```

## Response Variable by AMF with HEC fill

```{r}
#| fig-width: 12
#| fig-height: 8
plot_boxplot_cld(
  data         = ds_nu,
  response_var = "total_biomass_mg",
  x_var        = "amf",
  fill_var     = "hec",
  palette      = pal_hec,
  x_label      = "AMF Treatment",
  fill_label   = "HEC Treatment",
  y_label      = "Total Biomass (mg)",
  title        = "Total Biomass by AMF and HEC Treatment",
  ymax_offset  = 5                     # adjust letter height per variable
)

## Spad by AMF and HEC Treatment
#| fig-width: 12
#| fig-height: 8
plot_boxplot_cld(
  data         = ds_nu,
  response_var = "spad_avg",      
  x_var        = "amf",
  fill_var     = "hec",
  palette      = pal_hec,
  x_label      = "AMF Treatment",
  fill_label   = "HEC Treatment",
  y_label      = "Chlorophyl Content (SPAD)",
  title        = "Chlorophyl Content by AMF and HEC Treatment",
  ymax_offset  = 0.5   # <-- change this per chart as needed
)

```

## GR boxplot with hec on x-axis, amf as fill
```{r}
plot_boxplot_cld(
  data         = ds_nu,
  response_var = "gr",
  x_var        = "hec",
  fill_var     = "amf",
  palette      = pal_amf,
  x_label      = "HEC Treatment",
  fill_label   = "AMF Treatment",
  y_label      = "Plant Growth Response (GR)",
  title        = "Plant Growth Response to HEC and AMF Treatment",
  ymax_offset  = 0.1,
  caption_extra= "GR = log ( total biomass / x̄ total biomass of none HEC & sterile AMF)",
  hline        = 0
)

ggsave("C:\\Users\\sms689\\Downloads\\GR_boxplot.png", width = 8, height = 6, dpi = 300)
shell.exec("C:\\Users\\sms689\\Downloads\\GR_boxplot.png")
```

## AMF GR Boxplot HEC Fill
```{r}
#| fig-width: 12
#| fig-height: 8

ds_nu %>%
  ggplot(aes(x = hec, y = amf_gr, fill = hec)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.85, color = "gray30",
               stat = "summary",
               fun.data = function(x) {
                 m   <- mean(x, na.rm = TRUE)
                 sem <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
                 data.frame(
                   ymin   = min(x, na.rm = TRUE),  # whisker bottom
                   lower  = m - sem,
                   middle = m,
                   upper  = m + sem,
                   ymax   = max(x, na.rm = TRUE)   # whisker top
                 )
               }
  ) +
  geom_point(
    aes(color = hec),
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
    alpha = 0.5,
    size = 2
  ) +
  scale_fill_manual(name = "HEC Treatment", values = pal_hec) +
  scale_color_manual(name = "HEC Treatment", values = pal_hec) +
  theme_bw(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", hjust = 0.5, size = 18),
    axis.title       = element_text(face = "bold", size = 15),
    key              = element_blank(),
    axis.text.x      = element_text(size = 15),
    axis.text.y      = element_text(size = 15),
    panel.grid.minor = element_blank(),
    panel.border     = element_rect(color = "black", fill = NA),
    strip.text       = element_text(face = "bold")
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray30") +
  labs(
    y = "AMF Growth Response (GR)",
    x = "HEC Treatment",
    title = "AMF Plant Growth Response by HEC Treatment",
    subtitle = "Mean ± SEM, n = 3 per condition",
    caption = "AMF GR = (log (live AMF biomass / x̄ sterile AMF biomass))"
  )

ggsave("C:\\Users\\sms689\\Downloads\\AMF_GR_Boxplot_HEC_fill .bmp", width = 8, height = 12, dpi = 300)
```


## AMF GR Boxplot amf Fill
```{r}
#| fig-width: 12
#| fig-height: 8

ds_nu %>%
  ggplot(aes(x = hec, y = amf_gr, fill = amf)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.85, color = "gray30",
               stat = "summary",
               fun.data = function(x) {
                 m   <- mean(x, na.rm = TRUE)
                 sem <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
                 data.frame(
                   ymin   = min(x, na.rm = TRUE),  # whisker bottom
                   lower  = m - sem,
                   middle = m,
                   upper  = m + sem,
                   ymax   = max(x, na.rm = TRUE)   # whisker top
                 )
               }
  ) +
  geom_point(
    aes(color = amf),
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
    alpha = 0.5,
    size = 2
  ) +
  scale_fill_manual(name = "AMF Treatment", values = pal_amf) +
  scale_color_manual(name = "AMF Treatment", values = pal_amf) +
  theme_bw(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", hjust = 0.5, size = 18),
    axis.title       = element_text(face = "bold", size = 15),
    key              = element_blank(),
    axis.text.x      = element_text(size = 15),
    axis.text.y      = element_text(size = 15),
    panel.grid.minor = element_blank(),
    panel.border     = element_rect(color = "black", fill = NA),
    strip.text       = element_text(face = "bold")
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray30") +
  labs(
    y = "AMF Growth Response (GR)",
    x = "HEC Treatment",
    title = "AMF Plant Growth Response by HEC Treatment",
    subtitle = "Mean ± SEM, n = 3 per condition",
    caption = "AMF GR = (log (live AMF biomass / x̄ sterile AMF biomass))"
  )

ggsave("C:\\Users\\sms689\\Downloads\\AMF_GR_Boxplot_AMF_fill .bmp", width = 8, height = 12, dpi = 300)
```

## HEC GR Boxplot
```{r}
#| fig-width: 12
#| fig-height: 8
ds_nhn <- ds_nu %>%
  filter(hec != "none")
#View(ds_ns)
# 4. Plot
ds_nhn %>%
  ggplot(aes(x = hec, y = hec_gr, fill = amf)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.85, color = "gray30",
               stat = "summary",
               fun.data = function(x) {
                 m   <- mean(x, na.rm = TRUE)
                 sem <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
                 data.frame(
                   ymin   = min(x, na.rm = TRUE),  # whisker bottom
                   lower  = m - sem,
                   middle = m,
                   upper  = m + sem,
                   ymax   = max(x, na.rm = TRUE)   # whisker top
                 )
               }
  ) +
  geom_point(
    aes(color = amf),
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
    alpha = 0.5,
    size = 2
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray30") +
  scale_fill_manual(name = "AMF Treatment", values = pal_amf) +
  scale_color_manual(name = "AMF Treatment", values = pal_amf) +
  theme_bw(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", hjust = 0.5, size = 18),
    axis.title       = element_text(face = "bold", size = 15),
    axis.text.x      = element_text(size = 15),
    axis.text.y      = element_text(size = 15),
    panel.grid.minor = element_blank(),
    panel.border     = element_rect(color = "black", fill = NA),
    strip.text       = element_text(face = "bold")
  ) +
  labs(
    y = "HEC Growth Response (GR)",
    x = "HEC Treatment",
    title = "HEC Plant Growth Response by AMF Treatment",
    subtitle = "Mean ± SEM, n = 3 per condition",
    caption = "HEC GR = (log (HEC treatment biomass / x̄ none HEC biomass))"
  )
ggsave("C:\\Users\\sms689\\Downloads\\HEC_GR_Boxplot_amf_fill.bmp", width = 8, height = 12, dpi = 300)
```
