---
title: "EMF 28 Jan 2026"
editor: visual
format:
  html:
    code-fold: show
    code-summary: "Show the code"
author: 'Sedona Spann'
date: '28 Jan 2026'
---

# Background on Small Sample Sizes

Non-parametric Tests: When data violates normality assumptions or the sample is too small n\<15-20 , use Mann-Whitney U, Wilcoxon signed-rank, or Kruskal-Wallis tests.

Alternative Modeling

Bayesian Methods: Often more robust than frequent approaches for small samples.

Permutation Tests/Bootstrapping: Use resampling to estimate confidence intervals without assuming a normal distribution.

Generalized Linear Mixed Models (GLMMs): Use to account for nested data or random effects, which can improve power by structuring the error term.

Focus on Effect Sizes: Instead of just p-values, report effect sizes and confidence intervals to show the strength of findings rather than just significance.

Data Aggregation/Transformation:Convert annual data into quarterly or monthly data if appropriate.

Use the 10% rule: If the sample is less than 10% of the total population, observations can be treated as independent.

Reporting and Interpretation: Acknowledge low power and high risk of type II errors (false negatives). Emphasize that results are exploratory. For ecological modeling, ensure that interaction terms are kept to a minimum (pairwise) to avoid overfitting. 

# Hypotheses

**Hypothesis 1:**

**Hypothesis 2:**

# Install packages

```{r}
install.packages(c("tidyverse","lme4","sjPlot","ggsignif","emmeans","ggpubr","dplyr","ggplot2","lattice","glmmTMB","DHARMa","geepack","performance","ggeffects"))
```

# Load packages

```{r}
library(tidyverse)
library(lme4)
library(sjPlot)
library(ggsignif)
library(emmeans)
library(ggpubr)
library(dplyr)
library(ggplot2)
library(lattice)
library(glmmTMB)
library(DHARMa)
library(geepack)
library(performance)
library(ggeffects)
```

# Load data

```{r}
getwd()

# This is my path
setwd("C:/Users/sms689/Downloads")

ds <- read.csv("EMF Crop Growth Fall 2024.csv") # Using the cleaned up dataset.

View(ds)
head(ds)
```

## Calculate

```{r}
# Find the mean biomass of the dead pots for each site to use in MGR calculation
dead_means <- df %>%
  filter(Treatment=="ST") %>%
  group_by(Site, Seed_Source) %>%
  summarize(Mean_Dead_Shoot_Biomass_g=mean(Round_2_Dry_Shoot_Weight))

# Calculate SPAD averages
df <- df %>%
  mutate(SPAD_Avg = (SPAD_1 + SPAD_2 + SPAD_3) / 3)

# Calculate MGR using sterile plant biomass averages
df <- df %>%
  inner_join(dead_means, by=c("Site", "Seed_Source")) %>%
  mutate(MGR = case_when( # Calculate MGR
    Treatment != "ST" ~ log(Round_2_Dry_Shoot_Weight/Mean_Dead_Shoot_Biomass_g, base=exp(1)), # Apply the MGR formula to the live pots
    Treatment == "ST" ~ NA # Dead pots do not have an MGR value)
  ) )

df <- df %>%
  filter(Treatment != "ST")
```